---
title: "Assignment 5"
author: "Alex Ivina"
date: "November 26, 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}

library(tidyverse)
grad_enrollment<- read_csv("Grad_enrollment_1967 - 2015.csv")
library(effsize)
faculty_salary<- read_csv("faculty_salary.csv")
library(car)
library(kableExtra)
library(stargazer)
library(knitr)
library(reshape2)
```

1. Male and female graduate enrollment (1967 - 2015). Compare trends in total graduate enrollment for males and females (including full-time/part-time and private/public universities) in the United States from 1967 - 2015. Describe your results statistically, graphically and in text.

```{r}
#RESEARCH QUESTION: Does gender effect college enrollment rate


# Create a data set of male and female total enrollments
# look into regressions between genders
# Create a graphic

# data set of only totals for males and females
gender_totals<- grad_enrollment %>% 
  select(Year,Total,Total_Males, Total_Females)
gender_totals

#exploratory graphs
male_qq<- ggplot(gender_totals, aes(sample = Total_Males))+
  geom_qq()
male_qq
# data appears linear
male_hist<- ggplot(gender_totals, aes(x = Total_Males))+
  geom_histogram()
male_hist
# male enrollment appears relativly normally distributed


# Look into regression to see if there is a correlation between gender and college enrollment
gender_model_male<- lm(Total_Males~Year, data = gender_totals)
gender_model_male
# male_enrollment= -17112153 + 9069(Year)

# look to see if assumtions are met for regression
plot(gender_model_male)
# data look normal and linear the risiduals seem to be possibly 
summary(gender_model_male)
# R^2: 85 percent of male enrollment rate is determined by year
# p-value: variables are highly significant (p-value = 0)
AIC(gender_model_male)
# aic = 1210.979

#graph for male enrollment model
male_enrollment_graph<- ggplot(gender_totals, aes(y = Total_Males, x = Year))+
  geom_point()+
  geom_smooth(method = lm, se = TRUE, color = "chartreuse")+
  labs(y="Number of Students Enrolled", x = "Year")
male_enrollment_graph

######################################
# Look at the same qustion for women
#####################################

#exploratory graphs
female_qq<- ggplot(gender_totals, aes(sample = Total_Females))+
  geom_qq()
female_qq
# female enrollment data appeaes linear
female_hist<- ggplot(gender_totals, aes(x = Total_Females))+
  geom_histogram()
female_hist
# female data seem like it may not be normally distributed

#Look at a regression between year and female enrollment rate
female_model<- lm(Total_Females~Year, data = gender_totals)
female_model
# female_enrollment = -58955502 + 30126 (Year)

# look to see if assumtions are met for regression
plot(female_model)
# assumtions seem to be met, some worry with equal variances of risiduals
summary(female_model)
# R^2: 98% of female college enrollment rate is determined by year
# p-value: both variables are highly significant (p-vlaue = 0)
AIC(female_model)
# aic = 1217.443

# create a  final graphic for female enrollment rate

female_enrollment_graph<- ggplot(gender_totals, aes(x = Year, y = Total_Females))+
  geom_point()+
  geom_smooth(method = lm, se = TRUE, color = "orchid")+
  labs(y = "Number of Students Enrolled", x = "Year")
female_enrollment_graph


# Create one graph of males and females
enrollment_graph_0<- ggplot(gender_totals, aes(Year))+
  geom_line(aes(y = Total_Males, colour = "Total Males"))+
  geom_line(aes(y = Total_Females, colour = "Total Females"))+
  labs(title = "Male & Female College Enrollment U.S. (1967-2015)", x = "Year", y = "Number of Students Enrolled", colour = "Gender of Student")+
  scale_y_continuous(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0), limits = c(1967,2015))+
  theme_dark()

enrollment_graph_0

enrollment_graph<- ggplot(gender_totals, aes(Year))+
  geom_point(aes(y = Total_Males))+
  geom_smooth(method = lm, se = TRUE, size = 0.5, alpha = 0.5, aes(y = Total_Males, colour = "Total Males"))+
  geom_point(aes(y = Total_Females))+
  geom_smooth(method = lm, se = TRUE, size = 0.5, alpha = 0.5, aes(y = Total_Females, colour = "Total_Females"))+
  labs(x = "Year", y = "Number of Students", title = "Male and Female College Enrollment Rates", colour = "Gender of Student")
  
enrollment_graph

###Couple of things: Is there a way we can adjust the y-axis to include the female low of ~265,000 in 1967 and high of 172,000 in 2015? 

###Results need explanation: Good Example 1 (thorough):
#“Setosa sepal length significantly predicts sepal width (b = 0.79, t(48) = 7.6, p < 0.001) with a strong positive correlation between the two (Pearson’s r = 0.74). The overall model (width = 0.79(length) − 0.57; both in centimeters) explains a significant amount of variance in sepal width (F(1,48) = 59, p < 0.001, R2 = 0.55).”


###Something like: Year significantly predicts male and female graduate enrollment (statics reported respectfully). 

  

```


Written Results:
Total_Males-
Year significantly predicts college enrollment rates for males (b = 9069, t(47) = 16.61, p < 0.001) with a strong positive
correlation between the two (Pearson’s r = 0.92). The overall model (Total Male enrollment = -17,112,153 + 9069 (Year) explains a significant amount of variance in sepal width (F(1,47) = 476, p < 0.001, R2 = 0.8545).”

Total_Females-
Year significantly predicts college enrollment rates for females (b = 30126, t(47) = 51.66, p < 0.001) with a strong positive
correlation between the two (Pearson’s r = 0.99). The overall model (Total Female enrollment = -58955502 + 30126 (Year) explains a significant amount of variance in sepal width (F(1,47) = 2669, p < 0.001, R2 = 0.9827).



2. Shifts in female PhD recipients by field (1985, 2000, and 2015). Describe if and how there was a shift in PhDs awarded to females in four fields (Physical and Earth Sciences, Engineering, Education, and Humanities & Arts) in 1985, 2000, and 2015. Describe your results statistically, in a graph or table, and in text. Note: There are several ways that you can interpret this question. You are invited to decide which you think is/are most interesting. Just be really clear about what you are asking/answering in your report.

```{r}
Phds <- read_csv("Phds_female_fields.csv")

rownames(Phds) <- c("Female_Lifesciences", "Female_Physical_science", "Female_Math", "Female_Psychology", "Female_Engineering", "Female_Education", "Female_Humanities", "Female_Other") 

Phds_new <- select(Phds,"1985_Number":"2015_Number") %>% 
  rename("1985" = "1985_Number", "2000" = "2000_Number", "2015" = "2015_Number")


Phds_prop <- round(prop.table(as.matrix(Phds_new), 1), 2)

Phds_chi <- chisq.test(Phds_new)

Phds_chi

Phds_select <- subset(Phds_prop, rownames(Phds_prop) %in%
  c("Female_Physical_science","Female_Engineering","Female_Education","Female_Humanities"))

Phds_prop_table<- kable(Phds_select, col.names = c("1985","2000","2015"), caption = "Changes in field of study ") %>% 
  kable_styling(c("striped", "border",full_width = FALSE, position ="center"))
Phds_prop_table
  
Phds_2 <- as.data.frame(Phds_select) %>% 
  mutate(., Field = c("Physical & Eart Sciences","Engineering","Education","Arts & Humanities"))

Phds_2_expand <- melt(Phds_2,id.vars=c("Field"), value.name = "Proportion") 
names(Phds_2_expand)[2] <-paste("Year")

bar_graph_stack <- ggplot() +
  geom_bar(aes(y = Proportion, x = Year, fill = Field), data = Phds_2_expand, stat = "identity", width = 0.5) + coord_flip() + scale_fill_brewer(palette = 16) + labs(title = "Proportion of Female PhD Recipients by Field") +
   theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(expand=c(0,0))

bar_graph_stack



```

3. Male and female salaries for starting postdoctoral and other employment positions (2015). Compare median salaries for male and female doctorate recipients in 2015. Answer these two questions: Does median salary differ significantly between male and female starting postdoc positions? Does median salary differ significantly between male and female PhD recipients in non-postdoc employment positions?



=======
```{r,include=FALSE}

#read in the salaries dataframe from excel. Omit the extra "N/A" entries using the na.omit function
salary<-read_csv("Salaries.csv")%>%
  na.omit()
```

```{r, include=FALSE}
wsr_salary<-wilcox.test(salary$Job_male, salary$Job_female, paired =TRUE)


#There is a significant difference between male and female PhD recipients in non-postdoc employment positions in 2015 (V= 101, p=0.003).
```



```{r, include=FALSE}
wsr_salary
```

```{r, include=FALSE}
cliffsdelta<-cliff.delta(salary$Job_male, salary$Job_female, paired =TRUE)
cliffsdelta
```

```{r, include=FALSE}
wsr_salary_postdoc<-wilcox.test(salary$Postdoc_male, salary$postdoc_female, paired = TRUE)
wsr_salary_postdoc

#There is not a significant difference between male and female starting postdoctoral position salaries in 2015 (V=19.5, p=0.888).
```


```{r, include=FALSE}
wsr_salary_postdoc
```

```{r, include=FALSE}
cliffsdelta_postdoc<-cliff.delta(salary$Postdoc_male, salary$postdoc_female, paired = TRUE)
cliffsdelta_postdoc
```


Communicating Statistical Results:
Non-Parametric Wilcomxon Signed Rank anaylsis for data collected from 15 fields of study in the year 2015 revealed that the median salaries (in dollars) differed significantly for male and female PhD recipients in non-postdoc employment positions (V= 101, p=0.003). In all fields except Physics & Astonomy, men recieved either equal or moderately higher salaries(cliffs delta=0.213). 

However, a separate non-parametric wimcoxon signed rank test for PhD recipients starting posdoc positions revealed that there was *not* a signficant difference between salaries for males and females (V=19.5, p=.888). In several fields of study, males and females recieved the exact same salary. Females recieved a slightly higher median salary in some fields, while males were paid slightly more in others (cliffs delta=.04).



4. Exploring academic salaries for professors in U.S. colleges. Explore relationships between
variables in the ‘Faculty salary data (2008 - 2009 survey)’ dataset. Develop a model describing faculty
salary based on data for faculty sex, rank, years in current position, field, and number of years since
doctoral degree was earned. You should make decisions regarding which variables should remain in your
final model. Describe the results qualitatively and quantitatively (i.e., don’t just report the statistical
results of the model – make sure you describe interesting findings in text). You can also discuss any
concerns that you have with the model(s) you present, if any


```{r}


# data wrangling
faculty_salary1<- faculty_salary %>% 
  mutate(Discipline = case_when
         (`Discipline (A = Theoretical, B = Applied)` == "B" ~ "Apllied", 
                  `Discipline (A = Theoretical, B = Applied)`  == "A" ~"Theoretical")) %>% 
  select(-`Discipline (A = Theoretical, B = Applied)`) 

names(faculty_salary1)<- c("faculty_rank", "years_since_phd", "years_employed", "sex", "salary", "discipline")

faculty_salary1

# Exploratory graphs for salary
salary_qq<- ggplot(faculty_salary1, aes(sample = salary))+
  geom_qq()
salary_qq
# salary data looks linear
salary_hist<- ggplot(faculty_salary1, aes(x = salary))+
  geom_histogram()
salary_hist
# salary looks relatively normal

#################
# Create a model for salary
#################

salary_model<- lm(salary~faculty_rank + years_since_phd + years_employed + sex + discipline, data = faculty_salary1)
salary_model

# salary = 93280.4 + 535.1(years_since_phd) - 14417.6(discipline_theoretical) -12907.6(faculty_rankAsstProf) + 32158.4(faculty_rankProf) - 489.5(years_employed) + 4783.5(sexMale)

# look to see if regression assumptions are met
plot(salary_model)
# risidual seem to be normally distributed
summary(salary_model)
# all variables are significant EXCEPT sex, R^2 is relatively low (0.44) 
vif(salary_model)
# vif for all variables is under 4 so there is no colinearity between variables
AIC(salary_model)
# aic = 9093.826

############
# test to see if we can make the model better
###########

salary_model2<- lm(salary~faculty_rank + years_employed + sex + discipline, data = faculty_salary1)
salary_model2

# This model might not make sense because the coefficient for years_empoloyed is negative
summary(salary_model2)
AIC(salary_model2)
# aic = 9096.813
vif(salary_model2)
# this model doesnt have any variables that are highly correlated 

# try to take out years_employed and keep years_since_phd

salary_model3<- lm(salary~faculty_rank + years_since_phd  + sex + discipline, data = faculty_salary1)
salary_model3

summary(salary_model3)
AIC(salary_model3)
# aic = 9097.22
vif(salary_model3)
# This model also has no variables that are correlated

# explore the use of an interaction term between years_since_phd and years_employed

salarly_model4<- lm(salary~faculty_rank + years_since_phd + years_employed + sex + discipline + years_since_phd*years_employed, data = faculty_salary1)
salarly_model4

summary(salarly_model4)

AIC(salarly_model4)
# aic = 9088.57
vif(salarly_model4)
# super high correlations between years_since_phd, years_employed and the interaction term

####################
# make a model that does not include years_since_phd or years_employed
###################

salary_model5<- lm(salary~faculty_rank + sex + discipline, data = faculty_salary1)

summary(salary_model5)
vif(salary_model5) # no concern for colinearity
AIC(salary_model) # AIC = 9093.826

#################
# make a graph for faculty salary
#################

salary_graph_rank<- ggplot(faculty_salary1, aes(x = faculty_rank, y = salary))+
  geom_col()
salary_graph_rank

salary_graph_sex<- ggplot(faculty_salary1, aes(x = sex, y = salary))+
  geom_col()
salary_graph_sex

# think of ways to communicate multilinear regression, stargazer table??
# As of now use salary_model2 which exludes years_since_phd and includes years_employed



```

```{r, results = `asis`}

stargazer(salary_model2, type = "html")
# possibly use this graph when knitting to convay results not sure how to properly state results for multilinear regression like simple linevar regression. 
```








